### 2014-03-11 Abstract.tex
edited Abstract.tex

\section{Abstract}

Various MHD (magnetohydrodynamic) equilibrium tools, some of which being recently developped or considerably updated, are used on the medium-size COMPASS tokamak \cite{doi:10.1007/s10582-006-0188-1}. MHD equilibrium is a fundamental property of the tokamak plasma, whose knowledge is required for many diagnostics and modelling tools. Proper benchmarking and validation of equilibrium tools is thus key for interpretting and plannig tokamak experiments. We present here benchmarks and comparisons to experimental data of the EFIT++ reconstruction code [?], the free-boundary equilibrium code FREEBIE \cite{freebie2012}, and the rapid plasma boundary reconstruction part of the EQUINOX code \cite{doi:10.1016/j.jcp.2011.04.005}. We demonstrate that FREEBIE can calculate the equilibrium and corresponding poloidal field (PF) coils currents for given plasma parameters. Both EFIT++ and EQUINOX can reconstruct equilibria generated by FREEBIE from synthetic diagnostic data (including an artificial noise). Optimum reconstruction parameters are estimated; in addition, possible enhancements using more diagnostics are discussed and simulated using synthetic diagnostics. FREEBIE can also calculate the temporal evolution of the poloidal field coils currents for a whole plasma scenario.