### 2014-09-07 Introduction.tex
minor lean up

\section{Introduction} % (fold)
\label{sec:introduction}

This is intro

% section introduction (end)