### 2014-09-07 Procedure.tex

%!TEX root = JUrban_SOFT2014.tex

\section{Verification and validation procedure} % (fold)
\label{sec:procedure}

% section procedure (end)