### 2014-09-09 Results.tex
Example results for 6962 with noise

%!TEX root = JUrban_SOFT2014.tex

\section{Results} % (fold)
\label{sec:results}

We have selected five time slices from COMPAS shots 4275 and 6962 (i.e. 10 cases in total) for the analysis. These cases include circular, elongated and diverted plasmas with different currents. A comparison of plasma shapes for shot 4275 is shown in Fig. \ref{fig:ex4275}. Numerical values of reconstruction errors are presented in Table \ref{table:ex4275}. We can observe a very good agreement between the original equilibrium and the reconstructed shapes. In this case, FREEBIE was using linear $p'$ and $FF'$ polynomials so that the EFIT++ model agrees with the target data. VacTH uses 8 magnetic probes and 16 flux loops. As we discuss later, flux loops are essential for reliable VacTH results. Even global kinetic properties are well reconstructed in EFIT++; the largest error around 10~\% is in $l_{\mathrm i}$ (i.e. basically in the toroidal current density profile).

\begin{table*}
\centering

\begin{tabular}{lrrrrrrrrrrrrr}
\toprule
   code &  time &  $\Delta R_{\mathrm in}$ &  $\Delta R_{\mathrm out}$ &  $\Delta Z_{\mathrm min}$ &  $\Delta Z_{\mathrm max}$ &  $\delta \kappa$ &  $E_\mathrm{mp}$ &  $E_\mathrm{fl}$ &  $\delta W$ &  $\delta l_{\mathrm i}$ &  $\delta \beta_{\mathrm p}$ &  $\delta q_0$ &  $\delta q_{95}$ \\
\midrule
 EFIT++ &  0.97 &              0 &            0.002 &        0.001 &        0.001 &                  0.003 &       0.001 &      0.0009 &          0.04 &           0.09 &              0.03 &           0.03 &           0.007 \\
 EFIT++ &  0.99 &          4e-05 &            9e-05 &        0.001 &        0.001 &                  0.004 &       0.001 &      0.0006 &          0.04 &           0.09 &              0.04 &           0.02 &           0.005 \\
 EFIT++ &  1.02 &         0.0005 &           0.0002 &        0.002 &        0.002 &                  0.008 &       0.002 &       0.002 &          0.02 &            0.1 &              0.02 &           0.02 &           0.009 \\
 EFIT++ &  1.05 &          0.001 &           0.0008 &        4e-05 &       0.0003 &                  0.004 &       0.003 &       0.005 &          0.01 &           0.07 &              0.02 &           0.01 &           0.009 \\
 EFIT++ &   1.1 &          0.001 &           0.0005 &        0.005 &       0.0002 &                  0.005 &       0.004 &       0.002 &          0.04 &           0.09 &              0.03 &           0.02 &            0.05 \\
  VacTH &  0.97 &              0 &            0.001 &       0.0006 &       0.0002 &                  0.002 &       8e-07 &       7e-05 &           -- &            -- &               -- &            -- &             -- \\
  VacTH &  0.99 &          4e-05 &           0.0004 &        0.001 &       0.0008 &                  0.004 &       4e-07 &      0.0001 &           -- &            -- &               -- &            -- &             -- \\
  VacTH &  1.02 &          0.002 &           0.0008 &        0.005 &        0.002 &                   0.01 &       2e-06 &       0.002 &           -- &            -- &               -- &            -- &             -- \\
  VacTH &  1.05 &          0.006 &            0.002 &        5e-05 &        0.002 &                   0.01 &       3e-06 &        0.02 &           -- &            -- &               -- &            -- &             -- \\
  VacTH &   1.1 &          0.005 &            0.002 &        0.002 &        0.003 &                   0.02 &       2e-06 &       0.003 &           -- &            -- &               -- &            -- &             -- \\
\bottomrule
\end{tabular}

\caption{Errors for the same cases as in Fig. \ref{fig:ex4275}.}
\label{table:ex4275}
\end{table*}

\begin{figure*}
\centering   %\begin{center}
\hfill{}
\includegraphics[width=18cm]{figures/example_4275.pdf}
\hfill{}
%\end{center}
\caption{Contours of $\bar\psi=\left(0.25,0.5,0.75,1\right)$, reconstruction from FREEBIE data, shot 4275. EFIT++ parameters: $n_\mathrm{mp} = 16$, $n_\mathrm{fl} = 4$, $n_{p'} = n_{FF'} = 1$. VacTH parameters: $n_\mathrm{mp} = 8$, $n_\mathrm{fl} = 16$, $n_{\mathrm p} = n_{\mathrm q} = 5$.}
\label{fig:ex4275}
\end{figure*}

The second shot for the comparison is 6962, which has been chosen because Thomson scattering (TS) profiles are available. In Fig. \ref{fig:ex6962} we show reconstructions for more realistic equilibria and an artificial noise in the synthetic diagnostic signals. TS pressures were used in FREEBIE equilibria, which course no longer feature linear $p'$ and $FF'$ profiles. 
It is apparent that the reconstruction is not as good as in the previous case. Nevertheless, the plasma shape is well reconstructed with EFIT++ and VacTH, except for the last time slice in which VacTH yields too large plasma in the upper part.
$n_{\mathrm p} = n_{\mathrm q} = 4$ is used in this case as these values are minimum for reasonable VacTH results, while higher values are too sensitive to the input noise.
Expectedly, EFIT++ reconstruction with magnetic data only cannot reliably reconstruct kinetic plasma parameters, such as the stored energy or $q_0$. Rather unexpected is a relatively good agreement of $l_\mathrm{i}$. However, this agreement is compensated by a large error of $\beta_{\mathrm p}$ so that the quantity $\beta_{\mathrm p} + l_{\mathrm i}/2$ remain within a 10~\% error bar.

\begin{figure*}
\centering   %\begin{center}
\hfill{}
\includegraphics[width=18cm]{figures/example_6962_TS_noise.pdf}
\hfill{}
%\end{center}
\caption{Pressure profiles and contours of $\bar\psi=\left(0.25,0.5,0.75,1\right)$, reconstruction from FREEBIE data, shot 6962 with Thomson scattering pressure profiles. EFIT++ parameters: $n_\mathrm{mp} = 16$, $n_\mathrm{fl} = 4$, $n_{p'} = n_{FF'} = 1$. VacTH parameters: $n_\mathrm{mp} = 8$, $n_\mathrm{fl} = 16$, $n_{\mathrm p} = n_{\mathrm q} = 4$.}
\label{fig:ex6962}
\end{figure*}

\begin{table*}
\centering

\begin{tabular}{lrrrrrrrrrrrrr}
\toprule
   % code &  time &  $\Delta R_{\mathrm in}$ &  $\Delta R_{\mathrm out}$ &  $\Delta Z_{\mathrm min}$ &  $\Delta Z_{\mathrm max}$ &  $\delta \kappa$ &  $E_\mathrm{mp}$ &  $E_\mathrm{fl}$ &  $\delta W$ &  $\delta \left(\beta_{\mathrm p} + l_{\mathrm i}/2 \right)$  &  $\delta q_0$ &  $\delta q_{95}$ \\
   code &  time &  $\Delta R_{\mathrm in}$ &  $\Delta R_{\mathrm out}$ &  $\Delta Z_{\mathrm min}$ &  $\Delta Z_{\mathrm max}$ &  $\delta \kappa$ &  $E_\mathrm{mp}$ &  $E_\mathrm{fl}$ &  $\delta W$ &  $\delta l_{\mathrm i}$ &  $\delta \beta_{\mathrm p}$ &  $\delta q_0$ &  $\delta q_{95}$ \\
\midrule
 EFIT++ &  980 &               0 &            0.005 &        0.001 &        8e-06 &                   0.01 &        0.01 &        0.01 &           0.4 &           0.03 &               0.4 &            0.2 &          0.0008 \\
 EFIT++ & 1014 &           0.003 &            0.006 &         0.02 &         0.01 &                   0.06 &        0.03 &        0.03 &           0.3 &          0.008 &               0.3 &            0.2 &            0.08 \\
 EFIT++ & 1080 &           0.007 &            0.002 &        0.005 &        0.002 &                   0.04 &        0.03 &        0.03 &           0.3 &           0.01 &               0.3 &            0.2 &            0.02 \\
 EFIT++ & 1147 &           0.008 &            0.008 &         0.01 &        0.006 &                   0.06 &        0.03 &        0.02 &           0.3 &           0.06 &               0.3 &            0.3 &            0.05 \\
 EFIT++ & 1214 &          0.0002 &            0.008 &        0.002 &        0.001 &                   0.02 &        0.02 &        0.01 &          0.04 &            0.2 &              0.06 &            0.1 &            0.01 \\
  VacTH &  980 &               0 &             0.01 &         0.01 &        0.007 &                   0.02 &      0.0001 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
  VacTH & 1014 &           0.006 &           0.0002 &         0.01 &        0.007 &                   0.03 &       3e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
  VacTH & 1080 &           0.009 &           0.0007 &        0.003 &       0.0007 &                   0.02 &       3e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
  VacTH & 1147 &            0.02 &            0.007 &       0.0009 &        0.002 &                   0.01 &       5e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
  VacTH & 1214 &            0.03 &            0.009 &         0.02 &         0.03 &                   0.02 &      0.0002 &        0.02 &           -- &            -- &               -- &            -- &             -- \\

 % EFIT++ & 1e+03 &               0 &            0.005 &        0.001 &        8e-06 &                   0.01 &        0.01 &        0.01 &           0.4 &           0.03 &               0.4 &            0.2 &          0.0008 \\
 % EFIT++ & 1e+03 &           0.003 &            0.006 &         0.02 &         0.01 &                   0.06 &        0.03 &        0.03 &           0.3 &          0.008 &               0.3 &            0.2 &            0.08 \\
 % EFIT++ & 1e+03 &           0.007 &            0.002 &        0.005 &        0.002 &                   0.04 &        0.03 &        0.03 &           0.3 &           0.01 &               0.3 &            0.2 &            0.02 \\
 % EFIT++ & 1e+03 &           0.008 &            0.008 &         0.01 &        0.006 &                   0.06 &        0.03 &        0.02 &           0.3 &           0.06 &               0.3 &            0.3 &            0.05 \\
 % EFIT++ & 1e+03 &          0.0002 &            0.008 &        0.002 &        0.001 &                   0.02 &        0.02 &        0.01 &          0.04 &            0.2 &              0.06 &            0.1 &            0.01 \\
 %  VacTH & 1e+03 &               0 &             0.01 &         0.01 &        0.007 &                   0.02 &      0.0001 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
 %  VacTH & 1e+03 &           0.006 &           0.0002 &         0.01 &        0.007 &                   0.03 &       3e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
 %  VacTH & 1e+03 &           0.009 &           0.0007 &        0.003 &       0.0007 &                   0.02 &       3e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
 %  VacTH & 1e+03 &            0.02 &            0.007 &       0.0009 &        0.002 &                   0.01 &       5e-05 &        0.01 &           -- &            -- &               -- &            -- &             -- \\
 %  VacTH & 1e+03 &            0.03 &            0.009 &         0.02 &         0.03 &                   0.02 &      0.0002 &        0.02 &           -- &            -- &               -- &            -- &             -- \\

\bottomrule
\end{tabular}
\caption{Errors for the same cases as in Fig. \ref{fig:ex6962}.}
\label{table:ex6962}
\end{table*}

% section results (end)